### 2015-04-20 LaTeX/Bussmann14_text_WCv1p0.tex
Text from Cowley

We also compare our results to mock catalogues based on the methodology
presented in \cite{Cowley14}, for full details we refer the reader to Cowley et
al. but summarise it briefly here. An updated version of the \galform
\citep[e.g.][Lacey et al. in preparation]{Cole00} semi-analytic galaxy
formation model is used to populate halo merger trees
\citep[e.g.][]{PCH08,Jiang14} derived from a Millennium style $N$-body dark
matter only simulation \citep{Springel05,Guo13} with WMAP7 cosmology
\citep{Komatsu11}.  Submm flux is calculated using a model based on the
radiative transfer.  Dust is assumed to exist in two components,  dense
molecular clouds and a diffuse ISM.  Energy absorbed from starlight by each
dust component is calculated self-consistently. The dust is then assumed to
emit radiation as a modified blackbody, assuming thermal equilibrium.

Three randomly orientated 20 deg$^2$ lightcone catalogues are generated using
the method described in \cite{Merson13}.  We choose as the lower flux limit for
inclusion of simulated galaxies into our lightcone catalogue
$S_{500\mu\rm{m}}>0.1$ mJy, as this is the limit at which we recover 90 per
cent of the extragalactic background light (EBL) as predicted by our model (122
Jy deg$^{-2}$). This is in excellent agreement with observations from the
\emph{COBE} satellite \citep[e.g.][]{Puget96,Fixsen98}. This ensures we have a
realistic 500 $\mu$m  background in our mock images.

Mock imaging is created by binning the lightcone galaxies onto a pixelated grid
which is then convolved with a  36 arcsec FWHM Gaussian ($\sim$ \emph{Herschel}
SPIRE beam at 500 $\mu$m).  The image is then constrained to have a zero mean
by the subtraction of a uniform background.  For simplicity we do not attempt
to add any instrumental noise, nor perform any further filtering of the mock
image.

This procedure is repeated at 350 and 250 $\mu$m in order to provide simulated
\emph{Herschel} photometry.  We adjust the FWHM of the Gaussian PSF to 25 (18)
arcsec at 350 (250) $\mu$m, and change the lower limit of inclusion into our
lightcone to ensure 90 per cent of the predicted EBL is recovered at each
wavelength.

Source positions are then selected as maxima in the mock 250 $\mu$m image, with
the position and flux of the source being recorded as the center and value of
the maximal pixel.  To simulate `deblended' \emph{Herschel} photometry we
simply read off the value of the pixel located at the position of the $250$
$\mu$m maxima in the 350 and 500 $\mu$m images. For all \emph{Herschel} sources
with $S_{500\mu\rm{m}}>50$ mJy we then identify galaxies from our lightcone
catalogues within a 9 arcsec radius ($\sim$ ALMA primary beam) of the source
position, modelling the ALMA primary beam profile as a Gaussian with an 18
arcsec FWHM and a maximum sensitivity of 1 mJy.